BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications

### 2014-12-01 Paper/BOAST.tex
Improved Paper.

\section{Use Cases}
\label{use_cases}

  \subsection{Creating an Auto-Tuned Convolution Library for BigDFT using BOAST}
In 2005, the EU FP6-STREP-NEST BigDFT project funded a consortium of four European
laboratories (L\_Sim - CEA Grenoble, Basel University - Switzerland, Louvain-la-Neuve University - Belgium and
Kiel University - Germany), with the aim of developing a novel approach for DFT calculations
based on Daubechies wavelets.
Rather than simply building a DFT code from scratch, the objective of this three-years
project was to test the potential benefit of a new formalism in the context of electronic structure
calculations.

@@

% Convolutions are among the basic processing blocks of BigDFT. Special care has to be
% taken regarding their performances. 
The CPU convolutions of BigDFT have thus been thoroughly optimized.
In a recent paper, the optimization of the CPU convolutions of BigDFT has been extensively considered.
%BigDFT is characterized by its extensive use of
%convolution operators~\cite{nussbaumer1982fast} applied on large arrays of
%data.  
One example of a specific convolution, called
MagicFilter~\cite{Genovese|CAS2010}, can be seen in Listing~\ref{lst:conv_example}. 
It applies a filter $filt$ to the data set $in$ and then stores the result in the data set $out$ with a
transposition~\cite{Goedecker1993}.

@@

As we can see, there are three nested loops working on arrays whose sizes
vary.  Various optimizations can be applied to this treatment and
may focus on the loop structure, as well as on the size of the data. 
%In this
%paper we focus on this particular convolution and explore different
%optimizations whose nature is presented in detail in the next section.

    \subsubsection{A Generic Convolution Library}
The number of convolution kernels needed in BigDFT has been continuously growing in the recent years.
Various boundary conditions and functionalities have been added, making the BigDFT more and more powerful
in terms of scientific applications.
However, the cost of the maintainance and of development of the convolutions is always a 
delicate point to be considered while including a new functionality.
The convolution patterns are usually rather similar, leading to code duplication 
and difficulties in code reading and modification.

@@

The interest in having a robuts and optimally tuned library in this scientific field is therefore evident.
Techniques are under investigation to provide also fine-tuned binaries to the end-user rather than
the code sources, such that more agressive interprocedural optimisation can be performed.
This experience would be therefore a first step toward the release of a tunable optimised 
convolution library oriented to computational physics communities.
%   BigDFT implementation of convolution/wavelets is a reference but is limited in scope (1 wavelet family ands its derived operators).
%   Generalizing to other wavelet families long and difficult if performance are to remain good.
% 
% Idea:
% \begin{itemize}
% \item write a generic convolution library in BOAST
% \item BLAS like interface
% \item Provide a wide variety of building blocks
% \item Optimize thos blocks using BOAST
% \item Build either from source or directly link with binary
% \item Release complete library
% \end{itemize}
% 
% 
%   Characteristic use case driven
% 
%   Convolution Library: first step toward a dsl dedicated to wavelet processing for Physics computing.

>>>>

\section{BOAST Use Cases}
\label{use_cases}

In this section we will present two scientific applications that use BOAST. The
first one, BigDFT, uses boast in order to develop new functionalities with
performance portability in mind. The second one, SPECFEM3D, uses BOAST to
factorize OpenCL and CUDA development, while having robust non regression
tests.

  \subsection{Creating an Auto-Tuned Convolution Library for BigDFT using
BOAST}

In 2005, the EU FP6-STREP-NEST BigDFT project funded a consortium of
four European laboratories (L\_Sim - CEA Grenoble, Basel University -
Switzerland, Louvain-la-Neuve University - Belgium and Kiel University -
Germany), with the aim of developing a novel approach for DFT calculations
based on Daubechies wavelets.  Rather than simply building a DFT code from
scratch, the objective of this three-years project was to test the potential
benefit of a new formalism in the context of electronic structure calculations.

@@

% Convolutions are among the basic processing blocks of BigDFT. Special care
% has to be taken regarding their performances. 
The CPU convolutions of BigDFT have thus been thoroughly optimized.  In a
recent paper, the optimization of the CPU convolutions of BigDFT has been
extensively considered.
%BigDFT is characterized by its extensive use of convolution
%operators~\cite{nussbaumer1982fast} applied on large arrays of data.  
One example of a specific convolution, called
MagicFilter~\cite{Genovese|CAS2010}, can be seen in
Listing~\ref{lst:conv_example}.  It applies a filter $filt$ to the data set
$in$ and then stores the result in the data set $out$ with a
transposition~\cite{Goedecker1993}.

@@

As we can see, there are three nested loops working on arrays whose sizes vary.
Various optimizations can be applied to this treatment and may focus on the
loop structure, as well as on the size of the data. 
%In this paper we focus on this particular convolution and explore different
%optimizations whose nature is presented in detail in the next section.

    \subsubsection{A Generic Convolution Library}

The number of convolution kernels needed in BigDFT has been continuously
growing in the recent years.  Various boundary conditions and functionalities
have been added, making the BigDFT more and more powerful in terms of
scientific applications.  However, the cost of the maintainance and of
development of the convolutions is always a delicate point to be considered
while including a new functionality.  The convolution patterns are usually
rather similar, leading to code duplication and difficulties in code reading
and modification.

@@

The interest in having a robuts and optimally tuned library in this scientific
field is therefore evident.  Techniques are under investigation to provide also
fine-tuned binaries to the end-user rather than the code sources, such that
more agressive interprocedural optimisation can be performed.  This experience
would be therefore a first step toward the release of a tunable optimised
convolution library oriented to computational physics communities.
%   BigDFT implementation of convolution/wavelets is a reference but is limited
%   in scope (1 wavelet family ands its derived operators).  Generalizing to
%   other wavelet families long and difficult if performance are to remain
%   good.
% 
% Idea: \begin{itemize} \item write a generic convolution library in BOAST
% \item BLAS like interface \item Provide a wide variety of building blocks
% \item Optimize thos blocks using BOAST \item Build either from source or
% directly link with binary \item Release complete library \end{itemize}
% 
% 
%   Characteristic use case driven
% 
%   Convolution Library: first step toward a dsl dedicated to wavelet
%   processing for Physics computing.
